### 2020-11-26 Research Report - Boverhof/Tex report/main.tex
Multiple changes
- Changed structure of the repo
- Added latex table files
- Added the following parts to the report:
    * Data paragraph
    * Restructure of the report
    * Reformulation of several paragraphs

\usepackage[utf8]{inputenc}
\usepackage{apacite}
\usepackage{caption} 
\usepackage{setspace}

\captionsetup[table]{skip=4pt}
\onehalfspacing
\pagestyle{myheadings}

\begin{document}
\begin{titlepage}
\begin{center}
\LARGE{\textbf{The Realtime Assesment of Mental Workload by Means of Multiple Bio-Signals}}\\
\vspace*{2\baselineskip}
\Large{\textbf{Masterthesis Report}}\\
Methodology and Statistics for the Behavioural, Biomedical and Social Sciences\\
\vspace*{1\baselineskip}
Utrecht University\\
\vspace*{4\baselineskip}
{Bart-Jan Boverhof, 6000142}\\
\vspace*{1\baselineskip}
{\textbf{Thesis Supervisor}}\\
Prof.dr.ir. B.P. Veldkamp\\
\vspace*{1\baselineskip}
{\textbf{Date}}\\
October 14, 2020\\
\vspace*{1\baselineskip}
\end{center}
\end{titlepage}

\section{Introduction}
The topic of mental workload (MWL) has received widespread attention across a variety of different fields, amongst others the field of ergonomics \cite{young2015state}, human factors \cite{pretorius2007development} and neurosciences \cite{shuggi2017mental}.
A simple definition of MWL is the demand placed upon humans whilst carrying out a certain task. As pointed out by de Waard (1996), such a definition is too shallow, for it defines workload solely in external sense. It is of importance to acknowledge that MWL is person-specific, for the amount of experienced MWL ushered by a given task may differ across people \cite{de1996measurement}. When referring to MWL throughout this research, person-specific MWL is meant specifically.

@@

A real-time approach will be contrasted with a non-real-time approach. Special focus will be placed upon the justification of using a real-time approach, given the increment in complexity such an approach is likely to cause. Ultimately, this line of research pursues the ability to conduct a dynamic experiment for multiple people simultaneously, and which can be altered by the researches in real-time.

@@

\subsection{Related Work}
The current section will provide an overview of previous research on the most optimal model for each modality and the most feasible framework architectures.

\subsubsection{Electroencephalogram (EEG)}
The first utilized modality is a technique  called EEG, which detects electrical activity in the brain using electrodes. EEG is a widely utilized method for classifying MWL during experiments. With a review of the complete literature on EEG classification with deep learning, Craik and colleagues found a total of 16 \% of all available papers to deal with MWL, lending credence to the widely employed phenomenon of EEG for investigating MWL \cite{craik2019deep}. Additionally, this review aimed to map the feasibility of using deep neural learning for a range of different EEG application separately. Summarizing the findings, Craik and colleagues report that studies mostly found deep belief networks and ConvNets to perform best when aiming to classify MWL, and advice one of these approaches consequently.\cite{craik2019deep}

@@

\subsubsection{Galvanic Skin Response (GSR)}
The second utilized modality is GSR, measuring sweat gland on the hands and hereby inferring arousal. GSR readings have been found to significantly increase as a consequence of an increase in task cognitive load, hence constituting to be an objective predictor \cite{shi2007galvanic}.

@@

\subsubsection{Photoplethysmography (PPG)}
The third modality constitutes PPG, which is a technique utilized to measure heart rate. PPG is, not undeservedly, a widely deployed technique within the field of MWL classification. Zhang and colleagues investigated several approaches for measuring MWL, amongst three others the technique of PPG. Out of these four approaches, PPG was found to display both the highest sensitivity and reliability for measuring MWL, lending credence to the feasibility of PPG as a method for classifying MWL \cite{zhang2018evaluating}.

@@

\subsubsection{Framework Architecture - Data Fusion} 
When architecting a multi-model framework, it is of importance to realize that the information stemming from the different modalities are required to be combined (i.e. fused) in order to result in a single classification. Fusion can be done at different time-points within the framework. Several fusion strategies as proposed by Ramachandram and colleagues will be considered \cite{ramachandram2017deep}.

@@

\subsubsection{Framework Architecture - Real-time component}
bla bla

\subsection{Experimental Setup}
The experimental setting for data collection is the spaceship bridge simulator videogame Empty Epsilon, in which respondents are required to carry out tasks on a virtual spaceship bridge \cite{daid2016empty}. This experiment is instituted by the Brain Computer Interfaces (BCI) testbed lab hosted by the University of Twente (UT) and carried out in cooperation with Thales. The data is currently in the process of being collected. Ethical approval has been attained at the UT, and will be requested at the UU in the near future.\\  
\textit{Note on this paragraph: I will soon participate in the experiment, after which this paragraph will be extended.}

\subsection{Hyper Parameter Optimization}

\subsection{Model Evaluation}
The performance of the real-time framework will be assessed by contrasting it with a regular, non-real-time framework. Specifically, the quality of performance for each modality individually will be compared across both frameworks. This validation will be endeavored by means of six widely used performance measures, all constructed from the confusion matrix, as depicted below:

\bgroup
\def\arraystretch{1.6}%  
\begin{table}[h]
\centering
\caption{Confusion matrix}
\label{tab:my-table}
\begin{tabular}{lll}
                                        & True Positive          & True Negative          \\ \cline{2-3} 
\multicolumn{1}{l|}{Predicted Positive} & \multicolumn{1}{l|}{a} & \multicolumn{1}{l|}{b} \\ \cline{2-3} 
\multicolumn{1}{l|}{Predicted Negative} & \multicolumn{1}{l|}{c} & \multicolumn{1}{l|}{d} \\ \cline{2-3} 
\end{tabular}
\end{table}
\egroup

The measures accuracy, sensitivity, specificity, PPV, NPV and F1 will be utilized in order to asses model performance. The framework that performs best across these measures is considered to be the superior performing framework. Table two depicts the constitution of these performance measures by partly referring to confusion matrix depicted as table 1. 
\bgroup
\def\arraystretch{1.8}%  
\begin{table}[h]
\centering
\caption{Performance Metrics}
\label{tab:my-table}
\begin{tabular}{ll}
\hline
Accuracy:                       & \(\frac{\!\!\!\!\!\!\!\!\!\!\!\!\!\!a+d}{a+b+c+d}\) \\
Sensitivity:                    & \(\frac{a}{a+c}\)                                   \\
Specificity:                    & \(\frac{d}{b+d}\)                                   \\
Positive Predicted Value (PPV): & \(\frac{a}{a+b}\)                                   \\
Negative Predicted Value:       & \(\frac{d}{c+d}\)                                   \\
F1-measure:                     & \(\frac{2*Sensitivity*PPV}{Sensitivity+PPV}\)       \\ \hline
\end{tabular}
\end{table}
\egroup

>>>>

%Required packages
\usepackage[utf8]{inputenc}
\usepackage{apacite}
\usepackage{caption} 
\usepackage{setspace}

%Prerequisite statements
\captionsetup[table]{skip=4pt} %Line indent after table title
\captionsetup{labelfont={bf}} %Bold label font for tables
\onehalfspacing %Document line spacing
\pagestyle{myheadings}

\begin{document}

%Title page 
\begin{titlepage}
\begin{center}
\LARGE{\textbf{The Realtime Assesment of Mental Workload by Means of Multiple Bio-Signals}}\\
\vspace*{2\baselineskip}
\Large{\textbf{Masterthesis Report}}\\
Methodology and Statistics for the Behavioural, Biomedical and Social Sciences\\
\vspace*{1\baselineskip}
Utrecht University\\
\vspace*{4\baselineskip}
{Bart-Jan Boverhof, 6000142}\\
\vspace*{1\baselineskip}
{\textbf{Thesis Supervisor}}\\
Prof.dr.ir. B.P. Veldkamp\\
\vspace*{1\baselineskip}
{\textbf{Date}}\\
October 14, 2020\\
\vspace*{1\baselineskip}
\end{center}
\end{titlepage}

\section{Introduction} \label{Introduction}
The topic of mental workload (MWL) has received widespread attention across a variety of different fields, amongst others the field of ergonomics \cite{young2015state}, human factors \cite{pretorius2007development} and neurosciences \cite{shuggi2017mental}.
A simple definition of MWL is the demand placed upon humans whilst carrying out a certain task. As pointed out by de Waard (1996), such a definition is too shallow, for it defines workload solely in external sense. It is of importance to acknowledge that MWL is person-specific, for the amount of experienced MWL ushered by a given task may differ across people \cite{de1996measurement}. When referring to MWL throughout this research, person-specific MWL is meant specifically.

@@

A real-time classification approach will be contrasted with a non-real-time approach. Special focus will be placed upon the justification of using a real-time approach, given the increment in complexity such an approach is likely to cause. Ultimately, this line of research pursues the ability to conduct a dynamic experiment for multiple people simultaneously, and which can be altered by the researches in real-time.

@@

\subsection{Related Work} \label{Relatedwork}
The current section will provide an overview of previous research on the model design for each seperate modality. Additionally,  the most feasible architecture for a multi-modular framework will be explored. In particular, attention is placed upon the data fusion, the real-time component and several model optimization techniques.

\subsubsection{First modality: Electroencephalogram (EEG)}
The first utilized modality is a technique  called EEG, which detects electrical activity in the brain using electrodes. EEG is a widely utilized method for classifying MWL during experiments. With a review of the complete literature on EEG classification with deep learning, Craik and colleagues found a total of 16 \% of all available papers to deal with MWL, lending credence to the widely employed phenomenon of EEG for investigating MWL \cite{craik2019deep}. Additionally, this review aimed to map the feasibility of using deep neural learning for a range of different EEG application separately. Summarizing the findings, Craik and colleagues report that studies mostly found deep belief networks and ConvNets to perform best when aiming to classify MWL, and advice one of these approaches consequently.\cite{craik2019deep}

@@

\subsubsection{Second modality: Galvanic Skin Response (GSR)}
The second utilized modality is GSR, measuring sweat gland on the hands and hereby inferring arousal. GSR readings have been found to significantly increase as a consequence of an increase in task cognitive load, hence constituting to be an objective predictor \cite{shi2007galvanic}.

@@

\subsubsection{Third modality: Photoplethysmography (PPG)}
The third modality constitutes PPG, which is a technique utilized to measure heart rate. PPG is, not undeservedly, a widely deployed technique within the field of MWL classification. Zhang and colleagues investigated several approaches for measuring MWL, amongst three others the technique of PPG. Out of these four approaches, PPG was found to display both the highest sensitivity and reliability for measuring MWL, lending credence to the feasibility of PPG as a method for classifying MWL \cite{zhang2018evaluating}.

@@

\subsubsection{Framework Architecture}

\paragraph{Data fusion.}  
When architecting a multi-model framework, it is of importance to realize that the information stemming from the different modalities are required to be combined (i.e. fused) in order to ultimately result in a single classification. Fusion can be done at different time-points within the framework. Several fusion strategies as proposed by Ramachandram and colleagues will be considered \cite{ramachandram2017deep}.

@@

\paragraph{Real-time component}
bla bla

\subsection{Data}
The current section will provide an overview of the utilized data. Special attention is placed on the experimental setup, the description of the participants and the data collection / synchronization process.

\subsubsection{Experimental setup}
The experimental setting for data collection is the spaceship videogame Empty Epsilon, in which the respondent is required to carry out tasks on a virtual spaceship \cite{daid2016empty}. This experiment is instituted by the Brain Computer Interfaces (BCI) testbed lab, hosted by the University of Twente (UT) and carried out in cooperation with Thales. The experiment constituted three different segments, during each of which respondents had to carry out different tasks, all aiming to measure workload. Each segment consists of six small sessions of roughly 5-10 minutes. These sessions varied in difficulty, including two easy, two intermediate and two hard sessions per segment. A schematic overview of the experimental structure is depicted as table \ref{table:expsetup}. After each of the 18 segments, respondents filled in the TLX questionnaire consisting of 6 questions each, resulting in 18 filled in questionnaires. Each questionnaire inquired upon the degree to which the respondent experienced workload during the previous session. These ratings will be used as labels in later training. Within each segment, the order in which the sessions (varying in difficulty) were presented have been randomized. The order in which the segments were administrated were not random. Between every three sessions, respondents were requested to take a short 2 minute break.
\bgroup
\def\arraystretch{1.6}%  
\begin{table}[h]
\centering
\caption{Experimental setup: Number of sessions per segment and difficulty setting}
\label{table:expsetup}
\begin{tabular}{llll}
                               & Easy                   & Intermediate           & Hard                   \\ \cline{2-4} 
\multicolumn{1}{l|}{Segment 1} & \multicolumn{1}{l|}{2} & \multicolumn{1}{l|}{2} & \multicolumn{1}{l|}{2} \\ \cline{2-4} 
\multicolumn{1}{l|}{Segment 2} & \multicolumn{1}{l|}{2} & \multicolumn{1}{l|}{2} & \multicolumn{1}{l|}{2} \\ \cline{2-4} 
\multicolumn{1}{l|}{Segment 3} & \multicolumn{1}{l|}{2} & \multicolumn{1}{l|}{2} & \multicolumn{1}{l|}{2} \\ \cline{2-4} 
\end{tabular}
\end{table}
\egroup

The first segment emulated a scenario in which hostile spaceships approach the respondents spaceship. The respondent is required to  quickly react, and defuse hostile spaceships in order to survive. The increment in difficulty caused the process of defusing hostile spaceships to be more challenging and hereby to take longer,  aiming to increase workload consequently. The second segment emulated a scenario in which the respondent had to navigate their spaceship trough space, gathering as many way-points as possible. Obstacles around which the respondent had to carefully navigate were introduced in the intermediate difficult scenario, and hostile spaceships the respondent had to decimate were introduced in the hard scenario. Both increase difficulty, and hereby aim to increase workload consequently. The third and final segment emulated a machine room, in which respondents had to control the power based on random requests generated by the video-game. Variables that could overheat the spaceship  were introduced as a consequence of an increase in difficulty, demanding the respondent multi-task, hereby aiming to increase workload.

\subsubsection{Participants}
In total, \# respondents have participated in the study.

\subsubsection{Data collection and Synchronization}
About streaming to LSL, and the utilized devices.

\subsection{Framework}
The architecture for the proposed framework is constructed by combining insights from the literature, as delineated in paragraph \ref{Relatedwork} with the formulated design principles in section \ref{Introduction} (i.e. the principles of molecularity and generalizability).

The utilized network for the EEG modality will be a ConvNet with four convolutional max-pooling blocks. The utilized network for the GSR modality will be a hybrid LSTM ConvNet hybrid.

\subsection{Model Evaluation}
The performance of the real-time framework will be assessed by contrasting it with a regular, non-real-time framework. Specifically, the quality of performance for each modality individually will be compared across both frameworks. This validation will be endeavored by means of six widely used performance measures, all constructed from the confusion matrix, depicted as table  \ref{table:confusion}  
\bigskip
\bgroup
\def\arraystretch{1.6}%  
\begin{table}[h]
\centering
\caption{Confusion matrix}
\label{table:confusion}
\begin{tabular}{lll}
                                        & True Positive          & True Negative          \\ \cline{2-3} 
\multicolumn{1}{l|}{Predicted Positive} & \multicolumn{1}{l|}{a} & \multicolumn{1}{l|}{b} \\ \cline{2-3} 
\multicolumn{1}{l|}{Predicted Negative} & \multicolumn{1}{l|}{c} & \multicolumn{1}{l|}{d} \\ \cline{2-3} 
\end{tabular}
\end{table}
\egroup

The measures accuracy, sensitivity, specificity, PPV, NPV and F1 will be utilized in order to asses model performance. The framework that performs best across these measures is considered to be the superior performing framework. Table \ref{table:metrics} depicts the constitution of these performance measures by partly referring to confusion matrix depicted as table \ref{table:confusion}. 
\bigskip
\bgroup
\def\arraystretch{1.8}%  
\begin{table}[h]
\centering
\caption{Performance Metrics}
\label{table:metrics}
\begin{tabular}{ll}
\hline
Accuracy:                       & \(\frac{\!\!\!\!\!\!\!\!\!\!\!\!\!\!a+d}{a+b+c+d}\) \\
Sensitivity:                    & \(\frac{a}{a+c}\)                                   \\
Specificity:                    & \(\frac{d}{b+d}\)                                   \\
Positive Predicted Value (PPV): & \(\frac{a}{a+b}\)                                   \\
Negative Predicted Value:       & \(\frac{d}{c+d}\)                                   \\
F1-measure:                     & \(\frac{2*Sensitivity*PPV}{Sensitivity+PPV}\)       \\ \hline
\end{tabular}
\end{table}
\egroup
